# Project 3
## Megan Willis & Kyle Goulart

```{r, eval = FALSE}
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(ggthemes)
library(lubridate)
```


```{r}
# Our dataset: Spotify Songs
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```

Our Artists:
Tame Impala
A$AP Rocky
Fleetwood Mac
Queen
Childish Gambino
CCR
Elton John
Tyler, The Creator
Michael Jackson
Post Malone
Logic
Vampire Weekend
Chance
Mac DeMarco

### Filtering for our favorite artists
```{r}
favorite_artists = spotify_songs %>% 
  filter(track_artist %in% c("Tame Impala", "A$AP Rocky", "Fleetwood Mac", "Queen", "Childish Gambino", "Creedence Clearwater Revival", "Elton John", "Tyler, The Creator", "Michael Jackson", "Post Malone", "Logic", "Vampire Weekend", "Chance the Rapper", "Mac DeMarco")) %>% 
  select(track_artist, track_name, track_album_name, track_album_release_date, track_popularity, playlist_genre, danceability, energy, loudness, speechiness, acousticness, instrumentalness, liveness, tempo, duration_ms)
```

*Let's do some exploratory analysis*
```{r}
favorite_artists %>% 
  group_by(track_artist) %>% 
  count() %>% 
  ggplot(aes(x = fct_reorder(track_artist, n), y = n))+
  geom_col()+
  coord_flip()
```

**Talk about how the predictions may be more accurate for those with more songs**

**If predictive model is REALLY bad create a fct_lump other category for artists with small n**

**predicting genre by popularity?**

**release date and popularity**

### Let's create a column for release year

```{r}
favorite_artists = favorite_artists %>% 
  mutate(release_year = substring(track_album_release_date, 1, 4)) %>% 
  mutate(release_year = as.numeric(release_year))
```

